A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2021-02-10 01_TrialsForInclusionInAnalysis.Rmd
Rearrange trial inclusion and data selection for disease MA and Yield MA

Four trials from our subset need to be removed for not reporting yield variance.  
```{r TrialsWithVariance}
NoVar <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(
      location = location,
      year = year,
      No_Var = all(is.na(yield_error)),
      .groups = 'drop'
   ) %>%
   filter(No_Var) %>%
   distinct(trial_ref, year, location)
NoVar

PM_MB_dat <- PM_MB_dat %>%
   filter(!(trial_ref %in% NoVar$trial_ref))
```

@@

```{r Final_Trials_For_Inclusion}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   distinct(trial_ref, location, year) %>%
   arrange(year) 
```

Finally, we are going to define the class of each variable in the data, retaining only the variables relevant to the analysis. 
It should be noted that the data-set below has still retained all fungicide active ingredients. 
Before dropping the non-DMI fungicides from the final meta-analysis we will inspect their influence on grain yield.

```{r classifyData}
PM_MB_dat <-
   PM_MB_dat %>%
   mutate(trial_ref = as.character(trial_ref)) %>%
   mutate(year = as.character(year)) %>%
   mutate(location = as.character(location)) %>%
   mutate(host_genotype = as.character(host_genotype)) %>%
   mutate(row_spacing = as.double(round(row_spacing, digits = 2))) %>%
   mutate(replicates = as.integer(replicates)) %>%
   mutate(planting_date = ymd(planting_date)) %>%
   mutate(flowering_date = ymd(flowering_date)) %>%
   mutate(pod_fill_date = ymd(pod_fill_date)) %>%
   mutate(mid_late_pod_fill = ymd(mid_late_pod_fill)) %>%
   mutate(first_sign_disease = ymd(first_sign_disease)) %>%
   mutate(fungicide_ai = as.character(fungicide_ai)) %>%
   mutate(dose_ai.ha = round(as.double(dose_ai.ha), digits = 2)) %>%
   mutate(fungicide_application_1 = ymd(fungicide_application_1)) %>%
   mutate(fungicide_application_2 = ymd(fungicide_application_2)) %>%
   mutate(fungicide_application_3 = ymd(fungicide_application_3)) %>%
   mutate(fungicide_application_4 = ymd(fungicide_application_4)) %>%
   mutate(fungicide_application_5 = ymd(fungicide_application_5)) %>%
   mutate(fungicide_application_6 = ymd(fungicide_application_6)) %>%
   mutate(total_fungicide = as.integer(total_fungicide)) %>%
   mutate(harvest_date = ymd(harvest_date)) %>%
   mutate(final_assessment = ymd(final_assessment)) %>%
   mutate(PM_final_severity = round(as.double(PM_final_severity), digits = 4)) %>%
   mutate(disease_error = round(as.double(disease_error), digits = 6)) %>%
   mutate(D_error_type = as.character(D_error_type)) %>%
   mutate(grain_yield.t.ha. = round(as.double(grain_yield.t.ha.), digits = 4)) %>%
   mutate(yield_error = round(as.double(yield_error), digits = 6)) %>%
   mutate(Y_error_type = as.character(Y_error_type)) %>%
   mutate(Y_Msquare = round(as.double(Y_Msquare), digits = 6)) %>%
   mutate(yield_gain = round(as.double(yield_gain), digits = 4)) %>%
   mutate(prop_YG = round(as.double(prop_YG), digits = 4)) %>%
   select(
      -c(
         trial_design,
         # remove the following columns
         plot_length.m.,
         plot_width.m.,
         plant_density,
         trade_name,
         n_treatment,
         emergence_date,
         rating_scale,
         raw_graded,
         raw_data_avail,
         comments,
         days_harvest.planting,
         AUDPS_m,
         AUDPC_sd,
         Inc_Ms,
         pod_fill_date,
         mid_late_pod_fill,
         flowering_date,
         yield_gain,
         Y_Msquare,
         prop_YG,
         AUDPC_m,
         AUDPC_sd,
         AUDPS_sd
      )
   )
```

@@

@@

```{r clustered_fungicide_applications}
PM_MB_dat %<>%
   mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
   mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
   mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2)

@@

```{r simple_clusters}
PM_MB_dat <- PM_MB_dat %>%
   mutate(
      spray_management = case_when(
         fungicide_timing_1 < 0 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Recommended",
         fungicide_timing_1 >= 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Late",
         fungicide_timing_1 < 0 &
            !is.na(fungicide_application_2) ~ "Early_plus",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            !is.na(fungicide_application_2) ~ "Recommended_plus",
         fungicide_timing_1 >= 4 &
            !is.na(fungicide_application_2) ~ "Late_plus",
         TRUE ~ "Other"
      )
   )

@@

Now to view the number break-down of the `spray_management` treatments

```{r tableSprayManagement}
table(PM_MB_dat$spray_management)
```

@@

### Yield variance standardisation
```{r StandardiseYieldVariance}
PM_MB_dat <-
   PM_MB_dat %>%
   mutate(
      vi =
         case_when(
            Y_error_type == "stdev" ~ yield_error ^ 2,
            Y_error_type == "lsd (P=0.05)" ~
               (replicates * ((yield_error / 1.96) ^ 2) / 2) / replicates
         ),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   select(-c(yield_error,
             Y_error_type))

@@

### Disease severity variance standardisation
```{r StandardiseSeverityVariance}
PM_MB_dat <-
   PM_MB_dat %>%
   mutate(
      d_vi =
         case_when(
            D_error_type == "stdev" ~ disease_error ^ 2,
            D_error_type == "lsd (P=0.05)" ~
               (replicates * ((disease_error / 1.96) ^ 2) / 2) / replicates
         ),
      id = row_number(),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   select(-c(disease_error,
             D_error_type))

>>>>

@@

@@

## Excluding variance
Following an inspection of the yield and disease variance we determined imputation was not suitable and therefore we must exlude trials that did not report variance.
For each analysis, yield and disease severity a separate data-set will be created, including only trials that report variance for the respective response variable

Four trials will be removed for our yield data set `PM_dat_Y` for not reporting yield variance.  
```{r TrialsWithVariance}
NoVar <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(
      location = location,
      year = year,
      No_Var = all(is.na(yield_error)),
      .groups = 'drop'
   ) %>%
   distinct(trial_ref, year, location)
NoVar

PM_dat_Y <- PM_MB_dat %>%
   filter(!(trial_ref %in% NoVar$trial_ref))
```

Four trials will be removed for our disease severity data set `PM_dat_S` for not reporting yield variance.  
```{r TrialsWithVariance}
NoVar <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(
      location = location,
      year = year,
      No_Var = all(is.na(disease_error)),
      .groups = 'drop'
   ) %>%
   distinct(trial_ref, year, location)
NoVar

PM_dat_D <- PM_MB_dat %>%
   filter(!(trial_ref %in% NoVar$trial_ref))
```

```{r Final_Trials_For_Inclusion}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole") %>%
   mutate(MA_analysis = case_when(
      is.na(yield_error & disease_error) == FALSE ~ "Yield and Disease",
      is.na(yield_error) == FALSE & 
         is.na(disease_error) ~ "Yield",
      is.na(disease_error) == FALSE & 
         is.na(yield_error) ~ "Disease",
      TRUE ~ "Nil"
   )) %>%
   distinct(trial_ref, location, year, MA_analysis) %>%
   arrange(year) 
```

Finally, we are going to define the class of each variable in the data, retaining only the variables relevant to the analysis.

```{r classifyData}
source("R/classify_PM_data.R")

# classify yield data
PM_dat_Y <- 
   classify_PM_data(PM_dat_Y)

#classify disease data
PM_dat_D <- 
   classify_PM_data(PM_dat_D)
```

@@

```{r clustered_fungicide_applications}
# Yield data
PM_dat_Y %<>%
   mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
   mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
   mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2)

# Disease data
PM_dat_D %<>%
   mutate(fungicide_timing_1 = fungicide_application_1 - first_sign_disease) %>%
   mutate(fungicide_timing_2 = fungicide_application_2 - fungicide_application_1) %>%
   mutate(fungicide_timing_3 = fungicide_application_3 - fungicide_application_2)

@@

Simplify the clusters for the yield data-set
```{r simple_Y_clusters}
PM_dat_Y <- PM_dat_Y %>%
   mutate(
      spray_management = case_when(
         fungicide_timing_1 < 0 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Recommended",
         fungicide_timing_1 >= 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Late",
         fungicide_timing_1 < 0 &
            !is.na(fungicide_application_2) ~ "Early_plus",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            !is.na(fungicide_application_2) ~ "Recommended_plus",
         fungicide_timing_1 >= 4 &
            !is.na(fungicide_application_2) ~ "Late_plus",
         TRUE ~ "Other"
      )
   )

PM_dat_Y[PM_dat_Y$fungicide_ai == "control",
          c(
             "fungicide_timing_1",
             "fungicide_timing_2",
             "fungicide_timing_3",
             "spray_management"
          )] <- "control"
```

Simplify the clusters
```{r simple_D_clusters}
PM_dat_D <- PM_dat_D %>%
   mutate(
      spray_management = case_when(
         fungicide_timing_1 < 0 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Early",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Recommended",
         fungicide_timing_1 >= 4 &
            is.na(fungicide_application_2) &
            is.na(fungicide_application_3) ~ "Late",
         fungicide_timing_1 < 0 &
            !is.na(fungicide_application_2) ~ "Early_plus",
         fungicide_timing_1 >= 0 &
            fungicide_timing_1 < 4 &
            !is.na(fungicide_application_2) ~ "Recommended_plus",
         fungicide_timing_1 >= 4 &
            !is.na(fungicide_application_2) ~ "Late_plus",
         TRUE ~ "Other"
      )
   )

@@

Now to view the number break-down of the `spray_management` treatments in the yield and disease severity data-sets

```{r tableSprayManagement_Y}
table(PM_dat_Y$spray_management)
```

```{r tableSprayManagement_Y}
table(PM_dat_D$spray_management)
```

@@

### Yield variance standardisation
```{r StandardiseYieldVariance}
PM_dat_Y <-
   PM_dat_Y %>%
   mutate(
      vi =
         case_when(
            Y_error_type == "stdev" ~ yield_error ^ 2,
            Y_error_type == "lsd (P=0.05)" ~
               (replicates * ((yield_error / 1.96) ^ 2) / 2) / replicates
         ),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   select(-c(yield_error,
             Y_error_type))

@@

### Disease severity variance standardisation
```{r StandardiseSeverityVariance}
PM_dat_D <-
   PM_dat_D %>%
   mutate(
      d_vi =
         case_when(
            D_error_type == "stdev" ~ disease_error ^ 2,
            D_error_type == "lsd (P=0.05)" ~
               (replicates * ((disease_error / 1.96) ^ 2) / 2) / replicates
         ),
      id = row_number(),
      spray_management = fct_relevel(spray_management, sort)
   ) %>%
   select(-c(disease_error,
             D_error_type))
